---
title: "Preparing data for bnlearn models"
author: "Aaron Smith"
date: "6/24/2020"
output: word_document
---

Let's use bnlearn::preprocess() to discretize some data, then build Bayesian networks to compare the data preparation techniques.

The bnlearn::preprocess() function gives us three ways to discretize numeric data.

* interval: constant subinterval width
* quantile: all subintervals have (approximately) the same number of observations
* Hartemink: use mutual information to determine cutoffs

# Load the data

We will use the gaussian.test data from bnlearn.

```{r}
options(digits = 1)
require(bnlearn)

data(gaussian.test)
head(gaussian.test)
summary(gaussian.test)
```

# Discretize the data

Let's use the three discretize methods on the data set, and store the prepared data in a list.

```{r}
M <- gaussian.test

list_M <- lapply(
  X = c("interval","quantile","hartemink"),
  FUN = function(method) discretize(
    data = M,
    method = method,
    breaks = 4,
    ordered = TRUE
  )
)
names(list_M) <- c("interval","quantile","hartemink")
lapply(X = list_M,FUN = summary)
```

# Build the Bayesian networks

Now let's build the models. Some of the algorithms generate partially ordered networks. To make the models more consistent, let's use bnlearn::choose.direction() to order all graphs.

```{r}
v_algorithms <- c(
  "pc.stable","gs","iamb","fast.iamb","inter.iamb","iamb.fdr","mmpc","si.hiton.pc","hpc",
  "hc","tabu",
  "rsmax2","mmhc","h2pc",
  "aracne","chow.liu"
)
list_bnlearn <- list()
```

```{r,warning = FALSE}
for(j in v_algorithms) for(k in names(list_M)) try({
  list_bnlearn[[j]][[k]] <- do.call(
    what = j,
    args = list(x = list_M[[k]])
  )
  M_arcs <- arcs(list_bnlearn[[j]][[k]])
  for(l in 1:nrow(M_arcs)){
    list_bnlearn[[j]][[k]] <- set.arc(
      x = list_bnlearn[[j]][[k]],
      from = M_arcs[l,1],
      to = M_arcs[l,2],
      check.cycles = FALSE,
      check.illegal = FALSE
    )
    list_bnlearn[[j]][[k]] <- choose.direction(
      x = list_bnlearn[[j]][[k]],
      arc = M_arcs[l,],
      data = list_M[[k]]
    )
  }
},silent = TRUE)
```

# Scoring the networks

Now let's use Bayesian information criterion to evaluate each model.

```{r}
M_score <- matrix(
  data = NA,
  nrow = length(v_algorithms),
  ncol = length(list_M),
)
rownames(M_score) <- v_algorithms
colnames(M_score) <- names(list_M)

for(j in v_algorithms) for(k in names(list_M)) try({
  M_score[j,k] <- score(
    x = list_bnlearn[[j]][[k]],
    data = list_M[[k]],
    type = "bic"
  )
})
for(j in rownames(M_score)) M_score <- M_score[,order(M_score[j,])]
for(j in colnames(M_score)) M_score <- M_score[order(M_score[,j]),]
M_score
```

```{r}
graphviz.plot(
  list_bnlearn[[nrow(M_score)]][[ncol(M_score)]]
)
```